Community engagement and subgroup meta-knowledge: Some factors in the soul of a community

### 2015-08-05 PaperFinalDraft.Rnw
moving plus signs appropriately...

While the Gallup Poll analysis is interesting, it does raise the question of multicollinearity, as factors that are correlated with Community Attachment may simply be correlated with one of the variables that was used to compose it, and may not actually have an impact on local GDP growth. 
<<packages, echo=FALSE>>=
require(knitr)
@ 
<<chunkoptions, echo=FALSE, include=FALSE>>=
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.align='center')
require(dplyr)
require(ggplot2)
require(scales)
require(RColorBrewer)
require(ggmap)
require(xtable)
require(reshape2)
@
<<readdata>>=
sotc10 <- read.csv("data/sotc10.csv")
sotc09 <- read.csv("data/sotc09.csv")
sotc08 <- read.csv("data/sotc08.csv")
@
<<renaming>>=
renameData <- function(datasetname){
  plyr::rename(datasetname,c("INTRO1" = "gender", "QSB" = "citystate", "QS3"="county", "QS3_02"="county2", "QS3A"="zipcode", "QD1"="age", "QS4"="area_descrip", "QS5"="further", "QS5_2"="further2", "QN1A"="current_life", "QN1B"="future_life", "QCE1"="community_sat","QCE2"="recommend","Q3A"="proud", "Q3B"="perfect", "Q3C"="reputation", "Q4_1"="problem_1", "Q4_2"="problem_2", "Q4_3"="problem_3", "Q4_1_1"="problem_1", "Q4_1_2"="problem_2", "Q4_1_3"="problem_3","Q4_2_1"="problem_4", "Q4_2_2"="problem_5", "Q4_2_3"="problem_6","Q4_3_1"="problem_7", "Q4_3_2"="problem_8", "Q4_3_3"="problem_9","Q5"="rather_live", "Q6"= "five_years_ago", "Q6A" = "five_years_future", "Q7A"= "parks", "Q7B"= "beauty", "Q7C"= "highways", "Q7D"= "affordable_housing", "Q7E" =  "job_opportunites", "Q7F"=  "public_schools", "Q7G" = "colleges", "Q7H"= "nightlife", "Q7I"= "friends", "Q7J" = "raise_kids", "Q7K" = "heathcare", "Q7L" ="leadership", "Q7M" = "care", "Q7N"= "police", "Q7O" = "arts", "Q7P" =  "community_events", "Q8A" = "talented_grads", "Q8B" = "immigrants", "Q8C"= "minorities", "Q8D"= "families_kids", "Q8E" = "gay_lesbian", "Q8F" = "seniors", "Q8G" = "young_adults", "Q9" = "economy", "Q10" = "econ_better", "Q10A" = "trust_government", "Q11" = "employment", "Q12" = "commute", "Q13" = "job_satisfaction", "Q14" = "company_workforce", "Q15" = "enough_income", "Q15AA" = "area_hiring", "Q15AB" = "representative_leaders", "Q16A" = "treated_respectfully", "Q16B" = "well_rested", "Q16C" = "high_stress", "Q16D" = "learned_yest", "Q17" = "stress_source", "Q18" = "night_safety", "Q19" = "crime", "Q20" = "crime_increasing", "Q21"=  "registered_vote", "Q22A" = "volunteer", "Q22B" = "public_meeting", "Q22C"= "voted_local", "Q22D" = "worked_change", "Q22E" = "church", "Q22F" = "festival", "Q22G" = "donated_org", "Q22H" = "donated_individual", "Q22I" = "donated_shelter", "Q22_A" = "impact", "Q23" = "clubs", "Q24" = "friends_nearby", "Q24A" =  "friends_friends", "Q25"= "family_nearby", "Q26"= "talk_neighbors", "QD2" = "years_lived", "QD2A" ="permanent_resident", "QD3" = "work_cat", "QD3A" = "household_adults", "QD4" = "children", "QD5A" = "under6", "QD5B" = "six_12", "QD5C" = "thirteen_17", "QD6" = "married", "QD7"= "education", "QD8" = "own_home", "QD9"= "income", "QD10" = "hispanic", "QD111" = "race", "QD112" = "race2", "QD113" = "race3", "QD12" = "white_hispanic", "QD13" = "phone_numbers", "QD13A" = "cell_only", "QD13B" = "have_cell", "QD13C" = "cell_calls"))
}

@@

p <- ggplot(aes(x=year, y=percent), data=popData)
  + facet_wrap(~citystate, ncol=5)
  + geom_line()
  + scale_x_continuous(breaks=c(2008, 2008.5, 2009, 2009.5, 2010), 
                       labels=c("2008","", "2009", "", "2010"))
  + scale_y_log10(breaks=c(0.04, 0.4, 4))
  + xlab("")
  + ylab("Percent of community surveyed")
p
@
As mentioned above, the data were collected by Gallup through telephone surveys in 2008, 2009, and 2010. Participants were a random sample of adults living in 26 ``communities" (cities or metro areas of the United States). The data from 2008 and 2009 had \Sexpr{dim(sotc08)[1]} and \Sexpr{dim(sotc09)[1]} responses, respectively, while the data from 2010 contained \Sexpr{dim(sotc10)[1]} observations. Because the data sets surveyed the same 26 communities, we can calculate the average number of survey participants in each community. In 2008, that average was \Sexpr{floor(dim(sotc08)[1]/26)} people, in 2009, \Sexpr{floor(dim(sotc09)[1]/26)} people and in 2010, \Sexpr{floor(dim(sotc10)[1]/26)} people. The difference in average number of survey participants will be discussed further in Section \ref{missingdata}.

@@

<<communitysatplot, fig.cap="Responses to the question, \`\`Taking everything into account, how satisified are you with this community as a place to live?\" Communities are ordered by percentage of positive responses in 2008, making it clear the differences in distribution in 2009 and 2010.", out.width="0.99\\linewidth",fig.width = 9.5,fig.height=15,>>=
baseplot <- ggplot(mapping = aes(x=citystate, y=Freq, fill = Response, order=Response))
  + facet_wrap(~year, nrow=3)
  + geom_bar(data = trial2$neg, stat = "identity") 
  + scale_fill_manual(breaks=c("Not at all satisfied", "2", "3", "4", "Extremely satisfied"), 
                      values=colorsB, name="Response") 
  + geom_bar(data = trial2$pos, stat = "identity") 
  + coord_flip() 
  + ggtitle("Community satisfaction") 
  + xlab("") 
  + ylab("")
  + scale_y_continuous(limits=c(-0.5, 1), breaks=seq(from=-0.5, to=0.75, by=0.25), 
                       labels=c("50%", "25%", "0", "25%", "50%", "75%")) 
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text=element_text(size=14), strip.text=element_text(size=14))
baseplot 
@

@@

<<YNall, out.width="0.99\\linewidth",fig.width = 15,fig.height=9.5,fig.cap="Responses to yes/no questions about participants' behaviors, comparing all three survey years.">>=
baseplot <- ggplot(mapping = aes(Question, Freq, fill = Response))
  + facet_wrap(~Year)
  + geom_bar(data = trial1$neg, stat = "identity") 
  + geom_bar(data = trial1$pos, stat = "identity")
  + scale_y_continuous(breaks=seq(from=-0.75, to=0.75, by=0.25), labels=percent)
  + coord_flip()+ xlab("")
  + scale_x_discrete(labels=rev(c("Registered to vote \n", "Voted in local election \n", "Donated money to a \n local organization", "Attended a local \n festival or event", "Gave money or food \n to an individual", "Participated in a \n church event", "Performed local \n volunteer work", "Worked with other residents \n to make change", "Attended a local \n public meeting", "Provided free shelter \n to an individual")))
  + ylab("") 
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text.y=element_text(size=20), title=element_text(size=24), 
          axis.text.x=element_text(size=14), strip.text=element_text(size=18))
baseplot
@

@@

<<YN2010plot, out.width="0.99\\linewidth",fig.width = 10, fig.height=18, fig.cap="Percentage difference from overall survey rates (2010 data). The letters A-J represent the activities listed in Figure \\ref{fig:YNall}. A: Registered to vote, B: Voted in a local election, C: Donated money to a local organization, D: Attended a local event, E: Gave money or food to an individual, F: Participated in a church event, G: Performed local volunteer work, H: Worked with other residents to make change, I: Attended a local public meeting, J: Provided free shelter to an individual.">>=
baseplot <- ggplot(dfcitiesYes) 
  + aes(Question, diff, fill = col) 
  + facet_wrap(~citystate, ncol=3) 
  + geom_bar(stat = "identity")
  + coord_flip()
  + guides(fill=FALSE)
  + scale_y_continuous(breaks=seq(from=-.5, to=.5, by=.125),
                       labels=c("50%","", "25%","", "0","", "25%","", "50%"))
  +scale_x_discrete(labels=LETTERS[10:1])
  + ylab("") 
  + xlab("")
  + theme(axis.text.y=element_text(size=12), title=element_text(size=18), 
          axis.text.x=element_text(size=14), strip.text=element_text(size=16))
baseplot
@

@@

<<overallRaceplot, out.width="0.99\\linewidth",fig.width = 15,fig.height=10, fig.cap="Responses to the question, \`\`Which of these groups best describes your racial background?\" Notice that 2010 shows an overrepresentation of Refused responses, compared to the other years.">>=
raceOver <- ggplot(races) 
  + aes(race, n) 
  + geom_bar(stat="identity") 
  + facet_wrap(~year,nrow=3) 
  + ylab("Number of respondents")
  + xlab("")
  + theme(axis.text.y=element_text(size=20), title=element_text(size=24), 
          axis.text.x=element_text(size=18), strip.text=element_text(size=25))

raceOver + coord_flip()
@

@@

baseplot <- ggplot() 
  + aes(InGroup, Freq, fill = Response, order=Response)
  + facet_wrap(~year, nrow=3)
  + geom_bar(data = whiteYears$neg, stat = "identity") 
  + geom_bar(data = whiteYears$pos, stat = "identity") 
  + scale_fill_manual(breaks=c("Very bad", "2", "3", "4", "Very good"), values=colorsA, name="Response")
  + coord_flip()
# TRUE means minority, and in 2008 minorities were under-rating
  + scale_y_continuous(breaks=seq(from=-0.25, to=0.75, by=0.25), labels=percent)
  + scale_x_discrete(labels=c("White", "Non-white")) 
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text.y=element_text(size=16), title=element_text(size=16), 
          axis.text.x=element_text(size=14), strip.text=element_text(size=14))
  + xlab("") + ylab("")

@@

<<allMinorities,out.width="0.99\\linewidth",fig.width = 10.5, fig.height=12.5, fig.cap="Responses to the question, \`\`How is your community as a place for racial and ethnic minorities?\" faceted by community (2009 data)." >>=
baseplot <- ggplot() 
  + aes(InGroup, Freq, fill = Response, order=Response)
  + facet_wrap(~citystate, ncol=3)
  + geom_bar(data = minGroups09$neg, stat = "identity") 
  + geom_bar(data = minGroups09$pos, stat = "identity")
  + scale_fill_manual(values=colorsA, breaks=c("Very bad", "2", "3", "4", "Very good"),name="Response")
  + coord_flip()
#TRUE means minorities, in Aberdeen they are under-rating
  + scale_y_continuous(breaks=seq(from=-0.25, to=0.75, by=0.25), labels=percent)
  + scale_x_discrete(labels=c("White", "Non-white"))
  + xlab("") 
  + ylab("")
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text.y=element_text(size=16), title=element_text(size=16), 
          axis.text.x=element_text(size=14), strip.text=element_text(size=14))

@@

<<anotherlookplot, out.width="0.76\\linewidth", fig.cap=paste0("Relationship between positive responses to the question, \`\`How is your community as a place for minorities?\" comparing ratings of Whites and Non-whites. Each community is represented, and the plot uses 2009 data. Some communities, like State College, PA, are under-rated by Whites, some are over-rated, like Grand Forks, ND, and some are rated the same by both groups, like Gary, IN. The black line shows y=x, for comparison. Grey lines at x=", round(vert, digits=2), " and y=", round(horiz, digits=2), " show the mean ratings by each group.") >>=
al <- ggplot(newDF09)
  + geom_vline(aes(xintercept=vert), color="grey") 
  + geom_hline(aes(yintercept=horiz), color="grey") 
  + geom_abline(intercept=0, slope=1) 
  + geom_point(aes(y=white, x=nonwhite)) 
  + geom_text(data = newDF09[newDF09$citystate %in% c("Duluth, MN", "Gary, IN", "Grand Forks, ND", "State College, PA"),],
              aes(y=white, x=nonwhite, label = citystate), vjust=c(-0.6, -0.4, 1.3, -0.6), hjust=0, size=4) 
  + xlab("Non-whites rating") 
  + ylab("Whites rating") 
  + xlim(0.4, 1) 
  + ylim(0.4, 1) 
al
@

@@

baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order=Response))
  + facet_wrap(~year, nrow=3)
  + geom_bar(data = seniorYears2$neg, stat = "identity") 
  + geom_bar(data = seniorYears2$pos, stat = "identity") 
  + scale_fill_manual(breaks=c("Very bad", "2", "3", "4", "Very good"), values=colorsA, name="Response")
  + coord_flip()
  + scale_y_continuous(breaks=seq(from=-0.25, to=0.75, by=0.25), labels=percent)
  + scale_x_discrete(labels=c("Non-seniors", "Seniors"))
  + xlab(" ") 
  + ylab("")
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text.y=element_text(size=12), title=element_text(size=16), 
          axis.text.x=element_text(size=12), strip.text=element_text(size=14))

@@

<<seniorPlot, out.width="0.99\\linewidth",fig.width = 10.5, fig.height=12.5, fig.cap="Responses to the question, \`\`How is your community as a place for seniors?\" faceted by community (2009 data). Every community follows the pattern of over-rating by seniors, but some communities have a smaller discrepancy between ratings of seniors and non-seniors.">>=
baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order = Response)) 
  + facet_wrap(~citystate, ncol=3)
  + geom_bar(data = seniorGroup09$neg, stat = "identity") 
  + geom_bar(data = seniorGroup09$pos, stat = "identity") 
  + scale_fill_manual(breaks=c("Very bad", "2", "3", "4", "Very good"), values=colorsA, name="Response")
  + coord_flip()
  + scale_y_continuous(breaks=seq(from=-0.25, to=0.75, by=0.25), labels=percent)
  + scale_x_discrete(labels=c("Non-seniors", "Seniors"))
  + xlab("") 
  + ylab("")
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text.y=element_text(size=16), title=element_text(size=16), 
          axis.text.x=element_text(size=12),strip.text=element_text(size=14))

@@

<<anotherlookseniors, out.width="0.76\\linewidth", fig.cap=paste0("Relationship between positive responses to the question, \`\`How is your community as a place for seniors?\" comparing ratings of non-seniors and seniors. The black line shows y=x, for comparison. Grey lines at x=", round(mean(totPos$Seniors), digits=2), " and y=", round(mean(totPos$Nonseniors), digits=2), " show the mean ratings by each group.") >>=
sr <- ggplot(totPos)
  + geom_vline(aes(xintercept=vert), color="grey") 
  + geom_hline(aes(yintercept=horiz), color="grey") 
  + geom_abline(intercept=0, slope=1) 
  + geom_point(aes(y=Nonseniors, x=Seniors)) 
  + geom_text(data = totPos[totPos$citystate %in% c("Wichita, KS", "Gary, IN", "Bradenton, FL"),], 
              aes(y=Nonseniors, x=Seniors, label = citystate), vjust=c(-0.4,-0.4,1), hjust=c(1,0,0), size=4) 
  + xlab("Senior rating") + ylab("Non-senior rating") 
sr
@

@@

<<kidsplotOverall, out.width="0.99\\linewidth",fig.width = 8.5, fig.height=5.5, fig.cap="Responses to the question, \`\`How is your community as a place for families with young children?\" Families with children are defined as any household with children under the age of 18, Childless households are any households without children, whether or not the residents have adult children living elsewhere. Ratings were consistently similar between groups and across survey years.">>=
baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order=Response))
  + facet_wrap(~year, nrow=3)
  + geom_bar(data = kidYears$neg, stat = "identity") 
  + geom_bar(data = kidYears$pos, stat = "identity")
  + scale_fill_manual(breaks=c("Very bad", "2", "3", "4", "Very good"), values=colorsA, name="Response")
  + coord_flip()
  + scale_y_continuous(breaks=seq(from=-0.25, to=0.75, by=0.25), labels=percent)
  + scale_x_discrete(labels=c("Childless \n households", "Families \n with children"))
  + xlab("") 
  + ylab("") 
  + theme(legend.text=element_text(size=14), legend.title=element_text(size=16), 
          axis.text.y=element_text(size=16), title=element_text(size=16), 
          axis.text.x=element_text(size=14), strip.text=element_text(size=14))

>>>>

While the Gallup Poll analysis is interesting, it does raise the question of multicollinearity, as factors that are correlated with Community Attachment may simply be correlated with one of the variables that was used to compose it, and may not actually have an impact on local GDP growth. 
<<packages, echo=FALSE>>=
require(knitr)
@ 
<<chunkoptions, echo=FALSE, include=FALSE>>=
opts_chunk$set(echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.align='center')
require(dplyr)
require(ggplot2)
require(scales)
require(RColorBrewer)
require(ggmap)
require(xtable)
require(reshape2)
@
<<readdata>>=
sotc10 <- read.csv("data/sotc10.csv")
sotc09 <- read.csv("data/sotc09.csv")
sotc08 <- read.csv("data/sotc08.csv")
@
<<renaming>>=
renameData <- function(datasetname){
  plyr::rename(datasetname,c("INTRO1" = "gender", "QSB" = "citystate", "QS3"="county", "QS3_02"="county2", "QS3A"="zipcode", "QD1"="age", "QS4"="area_descrip", "QS5"="further", "QS5_2"="further2", "QN1A"="current_life", "QN1B"="future_life", "QCE1"="community_sat","QCE2"="recommend","Q3A"="proud", "Q3B"="perfect", "Q3C"="reputation", "Q4_1"="problem_1", "Q4_2"="problem_2", "Q4_3"="problem_3", "Q4_1_1"="problem_1", "Q4_1_2"="problem_2", "Q4_1_3"="problem_3","Q4_2_1"="problem_4", "Q4_2_2"="problem_5", "Q4_2_3"="problem_6","Q4_3_1"="problem_7", "Q4_3_2"="problem_8", "Q4_3_3"="problem_9","Q5"="rather_live", "Q6"= "five_years_ago", "Q6A" = "five_years_future", "Q7A"= "parks", "Q7B"= "beauty", "Q7C"= "highways", "Q7D"= "affordable_housing", "Q7E" =  "job_opportunites", "Q7F"=  "public_schools", "Q7G" = "colleges", "Q7H"= "nightlife", "Q7I"= "friends", "Q7J" = "raise_kids", "Q7K" = "heathcare", "Q7L" ="leadership", "Q7M" = "care", "Q7N"= "police", "Q7O" = "arts", "Q7P" =  "community_events", "Q8A" = "talented_grads", "Q8B" = "immigrants", "Q8C"= "minorities", "Q8D"= "families_kids", "Q8E" = "gay_lesbian", "Q8F" = "seniors", "Q8G" = "young_adults", "Q9" = "economy", "Q10" = "econ_better", "Q10A" = "trust_government", "Q11" = "employment", "Q12" = "commute", "Q13" = "job_satisfaction", "Q14" = "company_workforce", "Q15" = "enough_income", "Q15AA" = "area_hiring", "Q15AB" = "representative_leaders", "Q16A" = "treated_respectfully", "Q16B" = "well_rested", "Q16C" = "high_stress", "Q16D" = "learned_yest", "Q17" = "stress_source", "Q18" = "night_safety", "Q19" = "crime", "Q20" = "crime_increasing", "Q21"=  "registered_vote", "Q22A" = "volunteer", "Q22B" = "public_meeting", "Q22C"= "voted_local", "Q22D" = "worked_change", "Q22E" = "church", "Q22F" = "festival", "Q22G" = "donated_org", "Q22H" = "donated_individual", "Q22I" = "donated_shelter", "Q22_A" = "impact", "Q23" = "clubs", "Q24" = "friends_nearby", "Q24A" =  "friends_friends", "Q25"= "family_nearby", "Q26"= "talk_neighbors", "QD2" = "years_lived", "QD2A" ="permanent_resident", "QD3" = "work_cat", "QD3A" = "household_adults", "QD4" = "children", "QD5A" = "under6", "QD5B" = "six_12", "QD5C" = "thirteen_17", "QD6" = "married", "QD7"= "education", "QD8" = "own_home", "QD9"= "income", "QD10" = "hispanic", "QD111" = "race", "QD112" = "race2", "QD113" = "race3", "QD12" = "white_hispanic", "QD13" = "phone_numbers", "QD13A" = "cell_only", "QD13B" = "have_cell", "QD13C" = "cell_calls"))
}

@@

p <- ggplot(aes(x=year, y=percent), data=popData) + 
  facet_wrap(~citystate, ncol=5) + 
  geom_line() + 
  scale_x_continuous(
    breaks=c(2008, 2008.5, 2009, 2009.5, 2010), 
    labels=c("2008","", "2009", "", "2010")
    ) + 
  scale_y_log10(breaks=c(0.04, 0.4, 4)) + 
  xlab("") + 
  ylab("Percent of community surveyed")
p
@
As mentioned above, the data were collected by Gallup through telephone surveys in 2008, 2009, and 2010. Participants were a random sample of adults living in 26 ``communities" (cities or metro areas of the United States). The data from 2008 and 2009 had \Sexpr{dim(sotc08)[1]} and \Sexpr{dim(sotc09)[1]} responses, respectively, while the data from 2010 contained \Sexpr{dim(sotc10)[1]} observations. Because the data sets surveyed the same 26 communities, we can calculate the average number of survey participants in each community. In 2008, that average was \Sexpr{floor(dim(sotc08)[1]/26)} people, in 2009, \Sexpr{floor(dim(sotc09)[1]/26)} people and in 2010, \Sexpr{floor(dim(sotc10)[1]/26)} people. The difference in average number of survey participants will be discussed further in Section \ref{missingdata}.

@@

<<communitysatplot, fig.cap="Responses to the question, \`\`Taking everything into account, how satisified are you with this community as a place to live?\" Communities are ordered by percentage of positive responses in 2008, making it clear the differences in distribution in 2009 and 2010.", out.width="0.99\\linewidth",fig.width = 9.5,fig.height=15,>>=
baseplot <- ggplot(mapping = aes(x=citystate, y=Freq, fill = Response, order=Response)) + 
  facet_wrap(~year, nrow=3) + 
  geom_bar(data = trial2$neg, stat = "identity") + 
  scale_fill_manual(
    breaks=c("Not at all satisfied", "2", "3", "4", "Extremely satisfied"), 
    values=colorsB, 
    name="Response"
    ) + 
  geom_bar(data = trial2$pos, stat = "identity") + 
  coord_flip() + 
  ggtitle("Community satisfaction") + 
  xlab("") + 
  ylab("") + 
  scale_y_continuous(
    limits=c(-0.5, 1), 
    breaks=seq(from=-0.5, to=0.75, by=0.25), 
    labels=c("50%", "25%", "0", "25%", "50%", "75%")
    ) + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text=element_text(size=14), 
    strip.text=element_text(size=14))
baseplot 
@

@@

<<YNall, out.width="0.99\\linewidth",fig.width = 15,fig.height=9.5,fig.cap="Responses to yes/no questions about participants' behaviors, comparing all three survey years.">>=
baseplot <- ggplot(mapping = aes(Question, Freq, fill = Response)) + 
  facet_wrap(~Year) + 
  geom_bar(data = trial1$neg, stat = "identity") + 
  geom_bar(data = trial1$pos, stat = "identity") + 
  scale_y_continuous(
    breaks=seq(from=-0.75, to=0.75, by=0.25), 
    labels=percent
    ) + 
  coord_flip()+ 
  xlab("") + 
  scale_x_discrete(
    labels=rev(c("Registered to vote \n", "Voted in local election \n", "Donated money to a \n local organization", "Attended a local \n festival or event", "Gave money or food \n to an individual", "Participated in a \n church event", "Performed local \n volunteer work", "Worked with other residents \n to make change", "Attended a local \n public meeting", "Provided free shelter \n to an individual"))
    ) + 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=20), 
    title=element_text(size=24), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=18))
baseplot
@

@@

<<YN2010plot, out.width="0.99\\linewidth",fig.width = 10, fig.height=18, fig.cap="Percentage difference from overall survey rates (2010 data). The letters A-J represent the activities listed in Figure \\ref{fig:YNall}. A: Registered to vote, B: Voted in a local election, C: Donated money to a local organization, D: Attended a local event, E: Gave money or food to an individual, F: Participated in a church event, G: Performed local volunteer work, H: Worked with other residents to make change, I: Attended a local public meeting, J: Provided free shelter to an individual.">>=
baseplot <- ggplot(dfcitiesYes) + 
  aes(Question, diff, fill = col) + 
  facet_wrap(~citystate, ncol=3) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  guides(fill=FALSE) + 
  scale_y_continuous(
    breaks=seq(from=-.5, to=.5, by=.125),
    labels=c("50%","", "25%","", "0","", "25%","", "50%")
    ) +
  scale_x_discrete(labels=LETTERS[10:1]) + 
  ylab("") + 
  xlab("") + 
  theme(
    axis.text.y=element_text(size=12), 
    title=element_text(size=18), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=16))
baseplot
@

@@

<<overallRaceplot, out.width="0.99\\linewidth",fig.width = 15,fig.height=10, fig.cap="Responses to the question, \`\`Which of these groups best describes your racial background?\" Notice that 2010 shows an overrepresentation of Refused responses, compared to the other years.">>=
raceOver <- ggplot(races, mapping = aes(race, n)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~year,nrow=3) + 
  ylab("Number of respondents") + 
  xlab("") + 
  theme(
    axis.text.y=element_text(size=20), 
    title=element_text(size=24), 
    axis.text.x=element_text(size=18), 
    strip.text=element_text(size=25)) + 
  coord_flip()

raceOver 
@

@@

baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order=Response)) + 
  facet_wrap(~year, nrow=3) + 
  geom_bar(data = whiteYears$neg, stat = "identity")  + 
  geom_bar(data = whiteYears$pos, stat = "identity") + 
  scale_fill_manual(
    breaks=c("Very bad", "2", "3", "4", "Very good"), 
    values=colorsA, 
    name="Response"
    ) + 
  coord_flip() +
# TRUE means minority, and in 2008 minorities were under-rating
  scale_y_continuous(
    breaks=seq(from=-0.25, to=0.75, by=0.25), 
    labels=percent
    ) + 
  scale_x_discrete(labels=c("White", "Non-white"))  + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=16), 
    title=element_text(size=16), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=14)
    ) + 
  xlab("") + 
  ylab("")

@@

<<allMinorities,out.width="0.99\\linewidth",fig.width = 10.5, fig.height=12.5, fig.cap="Responses to the question, \`\`How is your community as a place for racial and ethnic minorities?\" faceted by community (2009 data)." >>=
baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order=Response)) + 
  facet_wrap(~citystate, ncol=3) + 
  geom_bar(data = minGroups09$neg, stat = "identity") + 
  geom_bar(data = minGroups09$pos, stat = "identity") + 
  scale_fill_manual(
    values=colorsA, 
    breaks=c("Very bad", "2", "3", "4", "Very good"),
    name="Response"
    ) + 
  coord_flip() +
#TRUE means minorities, in Aberdeen they are under-rating
  scale_y_continuous(
    breaks=seq(from=-0.25, to=0.75, by=0.25), 
    labels=percent
    ) + 
  scale_x_discrete(labels=c("White", "Non-white")) + 
  xlab("") + 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=16), 
    title=element_text(size=16), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=14))

@@

<<anotherlookplot, out.width="0.76\\linewidth", fig.cap=paste0("Relationship between positive responses to the question, \`\`How is your community as a place for minorities?\" comparing ratings of Whites and Non-whites. Each community is represented, and the plot uses 2009 data. Some communities, like State College, PA, are under-rated by Whites, some are over-rated, like Grand Forks, ND, and some are rated the same by both groups, like Gary, IN. The black line shows y=x, for comparison. Grey lines at x=", round(vert, digits=2), " and y=", round(horiz, digits=2), " show the mean ratings by each group.") >>=
al <- ggplot(newDF09) + 
  geom_vline(aes(xintercept=vert), color="grey") + 
  geom_hline(aes(yintercept=horiz), color="grey") + 
  geom_abline(intercept=0, slope=1) + 
  geom_point(aes(y=white, x=nonwhite)) + 
  geom_text(
    data = newDF09[newDF09$citystate %in% c("Duluth, MN", "Gary, IN", "Grand Forks, ND", "State College, PA"),],
    aes(y=white, x=nonwhite, label = citystate), 
    vjust=c(-0.6, -0.4, 1.3, -0.6), 
    hjust=0, 
    size=4
    ) + 
  xlab("Non-whites rating") + 
  ylab("Whites rating") + 
  xlim(0.4, 1) + 
  ylim(0.4, 1) 
al
@

@@

baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order=Response)) + 
  facet_wrap(~year, nrow=3) + 
  geom_bar(data = seniorYears2$neg, stat = "identity") + 
  geom_bar(data = seniorYears2$pos, stat = "identity") + 
  scale_fill_manual(breaks=c("Very bad", "2", "3", "4", "Very good"), values=colorsA, name="Response") + 
  coord_flip() + 
  scale_y_continuous(breaks=seq(from=-0.25, to=0.75, by=0.25), labels=percent) + 
  scale_x_discrete(labels=c("Non-seniors", "Seniors")) + 
  xlab(" ") + 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=12), 
    title=element_text(size=16), 
    axis.text.x=element_text(size=12), 
    strip.text=element_text(size=14))

@@

<<seniorPlot, out.width="0.99\\linewidth",fig.width = 10.5, fig.height=12.5, fig.cap="Responses to the question, \`\`How is your community as a place for seniors?\" faceted by community (2009 data). Every community follows the pattern of over-rating by seniors, but some communities have a smaller discrepancy between ratings of seniors and non-seniors.">>=
baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order = Response)) + 
  facet_wrap(~citystate, ncol=3) + 
  geom_bar(data = seniorGroup09$neg, stat = "identity") + 
  geom_bar(data = seniorGroup09$pos, stat = "identity") + 
  scale_fill_manual(
    breaks=c("Very bad", "2", "3", "4", "Very good"), 
    values=colorsA, 
    name="Response"
    ) + 
  coord_flip() + 
  scale_y_continuous(
    breaks=seq(from=-0.25, to=0.75, by=0.25), 
    labels=percent
    ) + 
  scale_x_discrete(labels=c("Non-seniors", "Seniors")) + 
  xlab("") + 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=16), 
    title=element_text(size=16), 
    axis.text.x=element_text(size=12),
    strip.text=element_text(size=14))

@@

<<anotherlookseniors, out.width="0.76\\linewidth", fig.cap=paste0("Relationship between positive responses to the question, \`\`How is your community as a place for seniors?\" comparing ratings of non-seniors and seniors. The black line shows y=x, for comparison. Grey lines at x=", round(mean(totPos$Seniors), digits=2), " and y=", round(mean(totPos$Nonseniors), digits=2), " show the mean ratings by each group.") >>=
sr <- ggplot(totPos) + 
  geom_vline(aes(xintercept=vert), color="grey") + 
  geom_hline(aes(yintercept=horiz), color="grey") + 
  geom_abline(intercept=0, slope=1) + 
  geom_point(aes(y=Nonseniors, x=Seniors)) + 
  geom_text(
    data = totPos[totPos$citystate %in% c("Wichita, KS", "Gary, IN", "Bradenton, FL"),], 
    aes(y=Nonseniors, x=Seniors, label = citystate), 
    vjust=c(-0.4,-0.4,1), 
    hjust=c(1,0,0), 
    size=4
    ) + 
  xlab("Senior rating") + 
  ylab("Non-senior rating")

sr
@

@@

<<kidsplotOverall, out.width="0.99\\linewidth",fig.width = 8.5, fig.height=5.5, fig.cap="Responses to the question, \`\`How is your community as a place for families with young children?\" Families with children are defined as any household with children under the age of 18, Childless households are any households without children, whether or not the residents have adult children living elsewhere. Ratings were consistently similar between groups and across survey years.">>=
baseplot <- ggplot(mapping = aes(InGroup, Freq, fill = Response, order=Response)) + 
  facet_wrap(~year, nrow=3) + 
  geom_bar(data = kidYears$neg, stat = "identity") + 
  geom_bar(data = kidYears$pos, stat = "identity") + 
  scale_fill_manual(
    breaks=c("Very bad", "2", "3", "4", "Very good"), 
    values=colorsA, 
    name="Response"
    ) + 
  coord_flip() + 
  scale_y_continuous(
    breaks=seq(from=-0.25, to=0.75, by=0.25), 
    labels=percent
    ) + 
  scale_x_discrete(labels=c("Childless \n households", "Families \n with children")) + 
  xlab("")+ 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=16), 
    title=element_text(size=16), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=14))
